---
title: "individual2"
author: "Felix Baez-Santiago"
date: "11/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab 5
# Drunk Driving and Fatal Accidents

*Loading Files*
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
#library(classdata)
library(lubridate)
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
person <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
```

2. Create a data frame containing the persons who are fatally hurt in the accidents (look up variable INJ_SEV)

```{r}
per2 <- person %>% filter(INJ_SEV == '4')
head(per2)
```

4. Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the most dangerous vehicle, and what you find from the map. (Hint: Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication. Use readxl::read_xlsx to read in the GLCs.)

```{r}
```


6. Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results.

```{r}
```


8. Is summer or winter more dangerous? Does this depend on states? Explore and explain.

```{r}
```