---
title: "individual2"
author: "Felix Baez-Santiago"
date: "11/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab 5
# Drunk Driving and Fatal Accidents

```{r}
acc <- read.csv('https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv', stringsAsFactors=FALSE)
per <- read.csv('https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv')
library(tidyverse)
library(lubridate)
```

2. Create a data frame containing the persons who are fatally hurt in the accidents (look up variable INJ_SEV)

```{r}
fatal <- per %>%
      filter(INJ_SEV == 4)
```
```{r}
mostDangerousMakes <- fatal %>%
      filter(!is.na(MAKE)) %>%
      group_by(STATE, MAKE) %>%
      summarize(n = n()) %>%
      group_by(STATE) %>%
      filter(n == max(n))

```
4. Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the most dangerous vehicle, and what you find from the map. (Hint: Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication. Use readxl::read_xlsx to read in the GLCs.)

```{r, dependson=c("load","severe","danger")}
#states <- map_data('state')
   # glcs <- readxl::read_xlsx('FRPP_GLC_-_United_StatesSep292020.xlsx') %>%
 #     mutate(`State Code`=as.integer(`State Code`)) %>%
 #     mutate(`County Code`=as.integer(`County Code`))
  #  make_labels <- glcs %>%
  #    distinct(`State Name`, `State Code`) %>%
   #   inner_join(mostDangerousMakes, by=c('State Code'='STATE')) %>%
   #   mutate(`State Name`=tolower(`State Name`)) %>%
   #   inner_join(states %>% group_by(region) %>% summarize(lat=(min(lat)+max(lat))/2,long=(min(long)+max(long))/2), by=c('State Name'='region')) %>%
    #  inner_join(read.csv('makeCodes.csv'))
   # ggplot(states, aes(x=long,y=lat,group=group)) + geom_path() + geom_text(data=make_labels, aes(x=long,y=lat,label=Model,group=NULL,label.size=0.2)) + coord_map() + ggtitle('Most Dangerous Make by State') + theme_void()
    
```



6. Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results.

```{r}
data %>%
      mutate(SEX=ifelse(SEX==1,'Male',ifelse(SEX==2,'Female','Unknown'))) %>%
      mutate(DAY_WEEK=wday(DAY_WEEK, label=TRUE)) %>%
      filter(HOUR <= 24) %>%
      group_by(DAY_WEEK, HOUR, SEX) %>%
      summarize(n=n()) %>%
      distinct(DAY_WEEK, HOUR, SEX, n) %>%
      ggplot(aes(x=HOUR,weight=n,fill=SEX)) + geom_bar(position='dodge') + facet_wrap(~DAY_WEEK) + ggtitle('Accidents by Day of Week, Hour, and Sex') + xlab('Hour of Day') + ylab('Number of Accidents')
   
```


8. Is summer or winter more dangerous? Does this depend on states? Explore and explain.

```{r}

```

  This shows that in the dataset there are more accidents in the Summer overall.
  
```{r}
accident <- countyStateCodes %>%
  select(StateCode, StateName) %>%
  right_join(accident, by = c("StateCode" = "STATE"))
byState <- accident %>% 
  filter(sumWint == 'Summer' | sumWint == 'Winter') %>%
  group_by(sumWint, StateName) %>%
  summarise(numAcc = n()) %>%
  arrange(desc(numAcc))
head(byState, 10)
```