---
title: "team"
author: "Kellan Bouwman"
date: "11/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab 5
# Drunk Driving and Fatal Accidents

*Loading Files*
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
library(classdata)
library(lubridate)

acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
person <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)

```

1. Are there some days of the week where more accidents happen than the others (see FARS manual, use variable DAY_WEEK)?

```{r}


accDays <- ggplot(acc, aes(DAY_WEEK)) + geom_bar(width=0.7) + theme_minimal()

print(accDays + labs(y="Count", x="Days of The Week", title = "Total Number of Accidents on each day of the Week"))
```

2. Create a data frame containing the persons who are fatally hurt in the accidents (look up variable INJ_SEV)

```{r}

```

3. Create a data frame containing the most dangerous vehicle make in each state. The number of persons fatally hit in the vehicle make is used to assess the (non-)safety of a make. Make sure to handle the missing values appropriately. (look up variable MAKE)

```{r}

```

4. Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the most dangerous vehicle, and what you find from the map. (Hint: Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication. Use readxl::read_xlsx to read in the GLCs.)

```{r}

```

5. Join the accident and person table (work out which variable(s) to use)

```{r}

```

6. Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results.

```{r}

```

7. Now plot a choropleth map of the number of deaths on a county level.

```{r}

```

8. Is summer or winter more dangerous? Does this depend on states? Explore and explain.

```{r}

```

