---
title: "team"
author: "Kellan Bouwman"
date: "11/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab 5
# Drunk Driving and Fatal Accidents

*Loading Files*
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
library(classdata)
library(lubridate)

acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)

makeCode <- read.csv("makeCodes.csv")
stateCode <- read.csv("stateCodes.csv")
```

1. Are there some days of the week where more accidents happen than the others (see FARS manual, use variable DAY_WEEK)?

```{r}
acc %>%
  mutate(recode(DAY_WEEK, "1" = "Sunday", "2" = "Monday", "3" = "Tuesday", "4" = "Wednesday", "5" = "Thursday", "6" = "Friday", "7"  = "Saturday")) -> mutAcc
print(mutAcc$DAY_WEEK)


accDays <- ggplot(acc, aes(DAY_WEEK)) + geom_bar(width=0.7) + theme_minimal()

print(accDays + labs(y="Count", x="Days of The Week", title = "Total Number of Accidents on each day of the Week"))
```

2. Create a data frame containing the persons who are fatally hurt in the accidents (look up variable INJ_SEV)

```{r}

```

3. Create a data frame containing the most dangerous vehicle make in each state. The number of persons fatally hit in the vehicle make is used to assess the (non-)safety of a make. Make sure to handle the missing values appropriately. (look up variable MAKE)

```{r}
makeCode %>%
  mutate(MAKE = Codes) -> makeCode

stateCode %>%
  mutate(STATE = Code) -> stateCode

per %>%
  full_join(makeCode) -> per
per %>%
  full_join(stateCode) -> per

fatal<- per %>%
  filter(INJ_SEV == 2)

danger<- fatal %>%
  group_by(State, Make) %>%
  summarize(n = n()) %>%
  group_by(State) %>%
  filter(n == max(n))

print(danger)
```

4. Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the most dangerous vehicle, and what you find from the map. (Hint: Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication. Use readxl::read_xlsx to read in the GLCs.)

```{r}

```

5. Join the accident and person table (work out which variable(s) to use)

   Joined by STATE, ST_CASE, VE_FORMS, COUNTY, DAY, MONTH, HOUR, MINUTE, RUR_URB, FUNC_SYS, HARM_E",MAN_COLL, SCH_BUS

```{r}
acc_per <- full_join(acc, per)
head(acc_per)
```

6. Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results.

```{r}

```

7. Now plot a choropleth map of the number of deaths on a county level.

```{r}
states <- full_join(acc_per, stateCode)
states %>%
  select(STATE, LONGITUD, LATITUDE) -> states


ggplot(fatal, aes(x = LONGITUD, y = LATITUDE, fill = n)) + 
  geom_polygon(aes(group = group)) + 
  geom_path(data = states, aes(x = LONGITUD, y = LATITUDE, group = group), inherit.aes=FALSE, size=2, color='white')
```

8. Is summer or winter more dangerous? Does this depend on states? Explore and explain.

```{r}

```

